### 2014-05-06 manuscript.tex
revised methods

\documentclass[review]{elsarticle}

\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
\usepackage{mathtools}
\usepackage{adjustbox}
\usepackage{natbib}
\usepackage{booktabs}
\modulolinenumbers[5]

\graphicspath{ {./figures/} }

\journal{IB200}

%% bibliography styles
%% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}

%%%%%%%%%%%%%%%%%%%%%%%
%% front matter
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Fossil-calibrated supermatrix phylogeny of Onagraceae: patterns of floral and chromosome evolution}

\author[berk]{William A. Freyman\corref{cor1}}
\ead{freyman@berkeley.edu}
\cortext[cor1]{Corresponding author}

\address[berk]{Jepson Herbarium and Department of Integrative Biology, University of California, Berkeley}

\begin{abstract}
blah blah
\end{abstract}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%
%% intro
%%%%%%%%%%%%%%%%%%%%%%%

\linenumbers

\section{Introduction}

Questions: are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
When did the major clades diverge?
What are the patterns of petal color and petal number evolution, and
are their evolution correlated?
What is the pattern of chromosome number evolution?

\section{Methods}

\paragraph{Supermatrix assembly}

Lythraceae was selected to use as an outgroup since previous molecular phylogenetic analyses
place it sister to Onagraceae \citep{conti1997, Sytsma2004}.
I downloaded all DNA sequences from GenBank release 200 PLN division and
performed an exhaustive all-by-all BLASTn \citep{blast} comparison of sequences in Onagraceae
and Lythraceae.
Using a BLASTn e-value of $1.0 \times 10^{-10}$ threshold and a sequence length
percent similarity cutoff of 0.5,
I constructed clusters of putative homologs using a single-linkage hierarchical clustering algorithm.
Subspecies names were removed from all sequences, and all but one sequence of each species was pruned from each cluster.
Clusters that were not phylogenetically informative ($< 4$ taxa) were discarded,
and each cluster was aligned using MUSCLE \citep{edgar2004muscle}. 
The alignments were concatenated by species, and any species that was not present in at least
two clusters was removed from the supermatrix.
The program written to data-mine GenBank and assemble the supermatrix 
is available as the open source Python module 
SUMAC \citep{sumac}.
SUMAC can be used to assemble
supermatrices for any taxonomic group recognized in GenBank, 
and is optimized to run on multicore processors and clusters by utilizing multiple parallel processes.

\paragraph{Phylogenetic analyses} 
Maximum likelihood (ML) analyses were performed with RAxML-HPC \citep{raxml} on the CIPRES Scientific Gateway \citep{cipres} 
using the rapid bootstrap heuristic and the GTRCAT nucleotide substitution model.
I used the ML tree to select 15 taxa phylogenetically widely distributed in Lythraceae to act as outgroup for the divergence time analysis; 
all other members of Lythraceae were subsequently removed from the supermatrix.
Bayesian estimates of divergence times were inferred using BEAST v1.8 \citep{beast, beast2} on CIPRES and calibrated with five fossils 
identified with morphological synapomorphies (Table \ref{fossils}).
The \textit{Ludwigia} fossil pollen was dated broadly to the Paleocene \citep{grimsson}, so I set the prior to a normal distribution with a wide 
standard deviation to cover the entire time period.
For all other calibration points I used a lognormal prior distribution with the offset (the minimum age of the node) corresponding to the fossil age.
The BEAST analysis utilized the GTR+$\Gamma$ nucleotide substitution model with a relaxed molecular clock (uncorrelated lognormal model)
and a Yule process tree prior.
The Markov Chain Monte Carlo (MCMC) was run for 100 million generations, sampling every 10 thousand generations.
Tracer v1.6 \citep{tracer} was used to assess the MCMC output for parameter convergence and ensure that the effective sample size for all parameters was above 200.
The first 1000 trees were discarded as burn-in, and the remaining 9000 trees were summarized as a maximum clade credibility (MCC) tree with mean divergence times.

\begin{table*}
   \center
   \begin{adjustbox}{max width=\textwidth}
      \begin{tabular}{lllllll}
         \toprule
         Group & Age (Mya) & Prior Distribution & Mean & SD & Offset & Reference \\ 
	 \midrule
         \textit{Circaea} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
         \textit{Epilobium} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
         S. Pacific \textit{Fuschia} (Onagraceae) & 23 & lognormal & 0.0 & 1.0 & 23 & \citep{lee2013fossil} \\
         \textit{Ludwigia} (Onagraceae) & Paleocene & normal & 60.0 & 3.0 & - & \citep{zhi} \\
         Lythraceae & 82 & lognormal & 0.0 & 2.0 & 82 & \citep{graham} \\
         \bottomrule
      \end{tabular}
   \end{adjustbox}
   \caption{Fossils used as priors in the Bayesian divergence time analysis.}
   \label{fossils}
\end{table*}

\paragraph{Character state reconstruction}
I scored six characters, including
chromosome number, floral merosity, petal color, and self-compatibility/incompatibility. 
Character data was assembled from the comprehensive \citet{wagner2007revised} Onagraceae monograph.
Ancestral character state reconstructions of petal number and petal color were performed using Mesquite v2.75 \citep{mesquite}
over the Bayesian MCC tree.
Characters were treated as unordered categorical data, and optimized using maximum likelihood
with the Markov k-state 1 parameter (Mk1) model \citep{lewis2001likelihood}.

%%%%%%%%%%%%%%%%%%%%%%%
%% results
%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}

\paragraph{Supermatrix assembly}
SUMAC evaluated 5571 Onagraceae and 2832 Lythraceae nucleotide sequences to construct the supermatrix. 
The completed supermatrix consisted of 11 clusters of homologous sequences (Table \ref{clusters}).
As used in the maximum likelihood analyses (before pruning the number of outgroup taxa), 
the supermatrix contained 521 taxa, was 31862 nucleotides long, and contained 93.0\% missing data.

\begin{table*}
   \center
   \begin{adjustbox}{max width=\textwidth}
      \begin{tabular}{lllllll}
         \toprule
         DNA Region & \# of Taxa & Aligned Length & Missing data (\%) & Taxon Coverage Density \\ 
	 \midrule
         ITS & 453 & 1746 & 13.2 & 0.87 \\
         trnL & 234 & 1429 & 55.2 & 0.45 \\
         rpl16 & 91 & 1414 & 82.6 & 0.17 \\
         rbcL & 77 & 1474 & 85.2 & 0.15 \\
         rps16 & 74 & 1016 & 85.8 & 0.14 \\
         rbcL & 64 & 1310 & 87.7 & 0.12 \\
         PgiC2 & 47 & 4028 & 91.0 & 0.09 \\
         matK & 37 & 921 & 92.9 & 0.07 \\
         ndhF & 37 & 2063 & 92.9 & 0.07 \\
         pgiC & 26 & 14709 & 95.0 & 0.05 \\
         R5 & 18 & 3129 & 96.6 & 0.03 \\
         \bottomrule
      \end{tabular}
   \end{adjustbox}
   \caption{Clusters of homologous sequences used to assemble the supermatrix.}
   \label{clusters}
\end{table*}

\paragraph{Phylogeny and divergence time estimates}
The topologies of the ML and Bayesian phylogenies were identical for all major clades within Onagraceae, 
so only the Bayesian MCC tree (Figures \ref{posteriors} and \ref{genera}) is shown here.
All Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities of $> 0.95$
except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31) (Figure \ref{posteriors}).
Onagraceae was found to diverge from Lythraceae at 109 Mya (Figure \ref{genera}). 
Divergence time estimates of other major clades and 95\% highest posterior density (HPD) intervals can be seen in Table \ref{times}.

\begin{table*}
   \center
   \begin{adjustbox}{max width=\textwidth}
      \begin{tabular}{lllllll}
         \toprule
         Clade & Mean Age (Mya) & 95\% HPD Min & 95\% HPD Max \\ 
	 \midrule
         Onagraceae / Lythraceae & 109 & 88 & 131 \\
         \textit{Ludwigia} & 97 & 76 & 118 \\
	 \textit{Hauya} & 49 & 35 & 64 \\
	 \textit{Circaea} / \textit{Fuchshia} & 37 & 28 & 47 \\
	 \textit{Lopezia} & 71 & 55 & 68 \\
	 \textit{Gongylocarpus} & 60 & 45 & 77 \\
	 \textit{Epilobium} & 49 & 38 & 60 \\
	 \textit{Chamerion} & 47 & 36 & 57 \\
	 \textit{Xylonagra} & 43 & 33 & 52 \\
	 \textit{Clarkia} & 40 & 32 & 48 \\
         \textit{Terapteron} & 19 & 10 & 29 \\
	 \textit{Camissoniopsis} / \textit{Neoholmgrenia} & 14 & 5 & 23 \\
	 \textit{Eremothera} / \textit{Camissonia} & 24 & 16 & 33 \\
	 \textit{Taraxia} & 30 & 22 & 38 \\
	 \textit{Chylismiella} / \textit{Gayophytum} & 20 & 10 & 30 \\
	 \textit{Eulobus} & 26 & 19 & 34 \\
	 \textit{Chylismia} / \textit{Oenothera} & 25 & 18 & 31 \\
         \bottomrule
      \end{tabular}
   \end{adjustbox}
   \caption{Bayesian divergence time estimates of major clades.}
   \label{times}
\end{table*}

\paragraph{Character evolution}
blah blah

\begin{table*}
   \center
   \begin{adjustbox}{max width=\textwidth}
      \setlength{\tabcolsep}{20pt}
      \begin{tabular}{lllllll}
         \toprule
	 & \multicolumn{4}{c}{Number of Petals} \\
         & 2 & 4 & 5 & 6 \\ 
	 \cmidrule{2-5}
	 Petal Color \\
         \hspace{3 mm} Pink & -0.005 & .011 & $ns$ & $ns$ \\
	 \hspace{3 mm} Yellow & -0.008 & .021 & $ns$ & $ns$ \\
	 \hspace{3 mm} White & 0.008 & 0.013 & -0.006 & -0.006 \\
	 \hspace{3 mm} Green & $ns$ & -0.011 & $ns$ & $ns$ \\
	 \hspace{3 mm} Red & $ns$ & $ns$ & $ns$ & $ns$ \\
         \bottomrule
      \end{tabular}
   \end{adjustbox}
   \caption{Test statistics for the correlation between flower color and number of petals. 
	    $d$ values are shown
            for the pairwise comparison of states with $p< 0.01$. $ns$ indicates no significant
	    association. The overall $D$ value was 0.263 ($p=0.00$).
	    }
   \label{correlations}
\end{table*}

\begin{figure*}[p]
    \vspace*{-2cm}
    \makebox[\linewidth]{
        % have to trim bottom off image
        \includegraphics[width=1.6\linewidth, trim=0 10 0 0, clip=true]{colored_posterior}
    }
    \caption{
       Bayesian maximum clade credibility phylogeny of 280 Onagraceae taxa and 15 Lythraceae taxa. 
       Estimated posterior probabilities close to 1.0 are shown in green. All genera described in
       \citet{wagner2007revised} were found to be monophyletic with posterior probabilities of $> 0.95$
       except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31).
    }
    \label{posteriors}
\end{figure*}

\begin{figure*}[p]
    \vspace*{-2cm}
    \makebox[\linewidth]{
        \includegraphics[width=1.6\linewidth, trim=0 10 0 0, clip=true]{time_colored_genera}
    }
    \caption{
       Bayesian chronogram of 280 Onagraceae taxa and 15 Lythraceae taxa.
       Approximate positions of fossil calibration points are shown as black circles.
       All genera described in \citet{wagner2007revised} are colored, and their
       divergence time estimates and \%95 HPD intervals can be seen in Table \ref{times}.
}
    \label{genera}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%
%% conclusion
%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

blah blah

%%%%%%%%%%%%%%%%%%%%%%%
%% references
%%%%%%%%%%%%%%%%%%%%%%%

\section*{References}

\bibliography{manuscriptbib}

\end{document}